PECK

### 2024-01-25 Data & Scripts/Alex Popescu - Final Stats.Rmd
Jan 25
Go over the script with a fine tooth comb:
Make sure the right labels are right
Change "behavior" to "behaviour"
Continue working on the Results and methods section, address comments
WITH THE NEW CHANGES, outline the discussion again

labs.bout.1<-c(
  '(Intercept)' = "Intercept"
  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
  , BEHAVIORHU = "Behavior"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
  , 'BEHAVIORHU:SENTINEL_PRESENCESentinel Absent' = "Behavior: Sentinel Presence"
  , 'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
)

labs.bout.2<-c(
  '(Intercept)' = "Intercept"
  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  ,'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
  )

@@

labs.ptwy<-c('(Intercept)' = "Intercept"
  , 'SENTINEL_PRESENCESentinel Absent' = "Sentinel Presence"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'SENTINEL_PRESENCESentinel Absent:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
  )
```

@@

```{r Open BOUT}
BOUT<-BOUT.raw %>%
  filter(.
         , DURATION > 0.01) %>% #Remove impossibly small values
  filter(.
         , BEHAVIOR != "M") %>% #Remove "M"
  mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
                         , 'YES' = "Sentinel Present"
                         , 'NO' = "Sentinel Absent"))
BOUT$LDURATION<-log(BOUT$DURATION) #Transform data for normality
str(BOUT)
```

@@

```{r BOUT Dot Plot}
BOUT.DOTPLOT<-BOUT.SUMMARY %>%
           ggplot(.
               , aes(x = GENERALIZED_ENVIRONMENT
                     , y = DURATION
                     , colour = SENTINEL_PRESENCE))+
  geom_point(position = position_dodge(width = 0.1)
             , size = 3
             ) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.1))+
  geom_text(aes(label = round(DURATION, 2))
            , hjust = -0.2
            , size = 6
            , show.legend = F
            , family = "serif") +
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , labels = c("Sentinel Absent", "Sentinel Present")
                      , name = "") +
  scale_x_discrete(labels = c("Commercial", "Green"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18, family = "serif")
        ) +
  facet_grid(~BEHAVIOR)

@@

```{r Subsetting PTWY}
PTWY<- DATA.SR %>%
  dplyr::select(VIDEO_ID,ID, DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_NB_TRANSITIONS, TOTAL_FREQUENCY_OF_DISTURBANCES | starts_with("NB_")) %>%
  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
         , "Transitions" = "TOTAL_NB_TRANSITIONS") %>%
  mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
                         , 'YES' = "Sentinel Present"
                         , 'NO' = "Sentinel Absent"))

@@

PTWY.BOXPLOT<-PTWY %>%
  rename("Foraging to Peck" = "NB_HD.HDP"
         , "Alert to Foraging" = "NB_HU.HD"
         , "Foraging to Alert" = "NB_HD.HU"
         , "Peck to Alert" = "NB_HDP.HU"
         ) %>%
  pivot_longer(., cols = c("Foraging to Peck"
                           , "Alert to Foraging"
                           , "Foraging to Alert"
                           , "Peck to Alert"
                           )
               , names_to = "TRANSITION"
               , values_to = "PTWY") %>%
    ggplot(.
           , aes(x = GENERALIZED_ENVIRONMENT
                 , y = PTWY
                 , colour = SENTINEL_PRESENCE)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(x = GENERALIZED_ENVIRONMENT
                  , group = SENTINEL_PRESENCE)
              , position = position_jitterdodge(jitter.width = 0.5)
              , alpha = 0.25
              ) +
  geom_point(data = PTWY.SUMMARY
             , aes( x = GENERALIZED_ENVIRONMENT
                    , y = PTWY
                    , group = SENTINEL_PRESENCE)
             , size = 2
             , colour = "black"
             , position = position_dodge(width = 0.75)) +
  geom_errorbar(data = PTWY.SUMMARY
                , aes(ymin=(PTWY-se)
                    , ymax=(PTWY+se)
                    , group = SENTINEL_PRESENCE
                    , x = GENERALIZED_ENVIRONMENT)
                , width = 0.2
                , colour = "black"
                , position = position_dodge(width=0.75
                                            )) +
  theme_classic() +
  ylab("Number of Transitions") +
  scale_colour_manual(values = cbPalette
                      , labels = c("Sentinel Absent"
                                   , "Sentinel Present")
                      , name = "") +
  scale_x_discrete(labels = c("Commercial", "Green")
                   )+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 14, family = "serif")
        ) +
  facet_grid(~TRANSITION)

>>>>

labs.bout.1<-c(
  '(Intercept)' = "Intercept"
  , 'SENTINEL_PRESENCESentinel Present' = "Sentinel Presence"
  , BEHAVIORHU = "Behavior"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'BEHAVIORHU:GENERALIZED_ENVIRONMENTGreen Area' = "Behavior: Generalized Environment"
  , 'BEHAVIORHU:SENTINEL_PRESENCESentinel Present' = "Behavior: Sentinel Presence"
  , 'SENTINEL_PRESENCESentinel Present:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
)

labs.bout.2<-c(
  '(Intercept)' = "Intercept"
  , 'SENTINEL_PRESENCESentinel Present' = "Sentinel Presence"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , GROUP_SIZESMALL = "Group Size"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  ,'SENTINEL_PRESENCESentinel Present:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
  )

@@

labs.ptwy<-c('(Intercept)' = "Intercept"
  , 'SENTINEL_PRESENCESentinel Present' = "Sentinel Presence"
  , 'GENERALIZED_ENVIRONMENTGreen Area' = "Generalized Environment"
  , BAIT_PRESENCEYES = "Bait Presence"
  , DISTURBANCE_FREQUENCY = "Disturbance Frequency"
  , 'SENTINEL_PRESENCESentinel Present:GENERALIZED_ENVIRONMENTGreen Area' = "Sentinel Presence: Generalized Environment"
  )
```

@@

```{r Open BOUT}
BOUT<-BOUT.raw %>%
  filter(.
         , DURATION > 0.01) %>% #Remove impossibly small values
  filter(.
         , BEHAVIOR != "M") %>% #Remove "M"
  mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
                                           , 'NO' = "Sentinel Absent"
                         , 'YES' = "Sentinel Present"
                         ))
BOUT$LDURATION<-log(BOUT$DURATION) #Transform data for normality
str(BOUT)
```

@@

```{r BOUT Dot Plot}
BOUT.DOTPLOT<-BOUT.SUMMARY %>%
           ggplot(.
               , aes(x = GENERALIZED_ENVIRONMENT
                     , y = DURATION
                     , colour = SENTINEL_PRESENCE))+
  geom_point(position = position_dodge(width = 0.1)
             , size = 3
             ) +
  geom_errorbar(aes(ymin=(DURATION-se)
                    , ymax=(DURATION+se))
                , width = 0.1
                , position = position_dodge(width=0.1))+
  geom_text(aes(label = round(DURATION, 2))
            , hjust = -0.2
            , size = 6
            , show.legend = F
            , family = "serif") +
  theme_classic() +
  ylab("Mean Bout Duration (s)") +
  scale_colour_manual(values = cbPalette
                      , name = "") +
  scale_x_discrete(labels = c("Commercial", "Green"))+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 18, family = "serif")
        ) +
  facet_grid(~BEHAVIOR)

@@

```{r Subsetting PTWY}
PTWY<- DATA.SR %>%
  dplyr::select(VIDEO_ID,ID, DECIMAL_TIME, GENERALIZED_ENVIRONMENT, SENTINEL_PRESENCE, BAIT_PRESENCE, NUMBER_OF_CROWS_RECORDED, GROUP_SIZE, TOTAL_NB_TRANSITIONS, TOTAL_FREQUENCY_OF_DISTURBANCES | starts_with("NB_")) %>%
  rename("DISTURBANCE_FREQUENCY" = "TOTAL_FREQUENCY_OF_DISTURBANCES"
         , "Transitions" = "TOTAL_NB_TRANSITIONS") %>%
  mutate(SENTINEL_PRESENCE = recode_factor(SENTINEL_PRESENCE
                                           , 'NO' = "Sentinel Absent"
                                           , 'YES' = "Sentinel Present"))

@@

PTWY.BOXPLOT<-PTWY %>%
  rename("Foraging to Peck" = "NB_HD.HDP"
         , "Alert to Foraging" = "NB_HU.HD"
         , "Foraging to Alert" = "NB_HD.HU"
         , "Peck to Alert" = "NB_HDP.HU"
         ) %>%
  pivot_longer(., cols = c("Foraging to Peck"
                           , "Alert to Foraging"
                           , "Foraging to Alert"
                           , "Peck to Alert"
                           )
               , names_to = "TRANSITION"
               , values_to = "PTWY") %>%
    ggplot(.
           , aes(x = GENERALIZED_ENVIRONMENT
                 , y = PTWY
                 , colour = SENTINEL_PRESENCE)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(aes(x = GENERALIZED_ENVIRONMENT
                  , group = SENTINEL_PRESENCE)
              , position = position_jitterdodge(jitter.width = 0.5)
              , alpha = 0.25
              ) +
  geom_point(data = PTWY.SUMMARY
             , aes( x = GENERALIZED_ENVIRONMENT
                    , y = PTWY
                    , group = SENTINEL_PRESENCE)
             , size = 2
             , colour = "black"
             , position = position_dodge(width = 0.75)) +
  geom_errorbar(data = PTWY.SUMMARY
                , aes(ymin=(PTWY-se)
                    , ymax=(PTWY+se)
                    , group = SENTINEL_PRESENCE
                    , x = GENERALIZED_ENVIRONMENT)
                , width = 0.2
                , colour = "black"
                , position = position_dodge(width=0.75
                                            )) +
  theme_classic() +
  ylab("Number of Transitions") +
  ylim(0,100)+
  scale_colour_manual(values = cbPalette
                      , name = "") +
  scale_x_discrete(labels = c("Commercial", "Green")
                   )+
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin()
        , text = element_text(size = 14, family = "serif")
        ) +
  facet_grid(~TRANSITION)

### 2024-01-25 Data & Scripts/PECK.Rmd
Jan 25
Go over the script with a fine tooth comb:
Make sure the right labels are right
Change "behavior" to "behaviour"
Continue working on the Results and methods section, address comments
WITH THE NEW CHANGES, outline the discussion again

```{r Peck Summary, echo = F}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
PECK.MEAN<-summarySE(DATA, "PECK_RATE", c("GENERALIZED_ENVIRONMENT", "SENTINEL_PRESENCE"), na.rm=T)
PECK.MEAN
```

@@

VERY WEIRD. Model 1 without the interactions is preferred over Model 2. However, the simple MOD0 is preferred. Let's see the outputs of that model.

```{r Robust PECK MODEL 0}
sjPlot::tab_model(PECK.MOD0
                  , show.re.var = T
                  , title = "Robust Peck Rate Model 0 Output"
                  , dv.labels = " Effects on peck rate")
```

@@

```{r PECK Disturb 1}
PK.DF.DOT1<-DATA %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE))+
  geom_point(size = 3) +
  theme_classic() +
  labs(y="Pecks per minute", x="Disturbances per Minute")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom")

@@

```{r PECK Disturb 2}
PK.DF.DOT2<-DATA[which(DATA$BEHAVIOR == "HD" & DATA$BEHAVIOR_DURATION > 0),] %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE)) +
  geom_point(size = 3) +
  geom_smooth(method="lm") +
  theme_classic() +
  labs(y="Pecks per minute", x="Disturbances per Minute")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom")

@@

Very nice! While it is not the prettiest of results, it roughly appears that as the disturbance frequency increases, so does the peck rate. I believe this can be explained by a sense of 'urgency' in higher-risk environments. Greater sample sizes would be required to make conclusive statements.

>>>>

```{r Peck Summary, echo = F}
### Summarize
PECK.MEAN<-summarySE(DATA, "PECK_RATE", c("GENERALIZED_ENVIRONMENT", "SENTINEL_PRESENCE"), na.rm=T)
PECK.MEAN
```

@@

VERY WEIRD. Model 1 without the interactions is perferred over Model 2. However, the simple MOD0 is preferred. Let's see the outputs of that model.

```{r Robust PECK MODEL 2}
sjPlot::tab_model(PECK.MOD0
                  , show.re.var = T
                  , title = "Robust Peck Rate Model 0 Output"
                  , dv.labels = " Effects on peck rate")
```

@@

```{r PECK Disturb}
PK.DF.DOT1<-DATA %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE))+
  geom_point(size = 3) +
  theme_classic() +
  labs(y="Pecks per minute", x="Disturbances per Minute")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom")

@@

```{r PECK Disturb}
PK.DF.DOT2<-DATA[which(DATA$BEHAVIOR == "HD" & DATA$BEHAVIOR_DURATION > 0),] %>%
    ggplot(aes(x = DISTURBANCE_FREQUENCY
               , y = PECK_RATE)) +
  geom_point(size = 3) +
  geom_smooth(method="lm") +
  theme_classic() +
  labs(y="Pecks per minute", x="Disturbances per Minute")+
  scale_x_continuous(n.breaks=14)+
  theme(legend.position = "bottom")

@@

Very nice! We can see that as the disturbance frequency increases, so does the peck rate. I believe this can be explained by a sense of 'urgency' in higher-risk environments.
